---
title: "figure s2"
output: html_notebook
---

Analyze effect of *per* knockout on both seasonal (diapause incidence) AND daily (eclosion) timing of *O. nubilalis*, as visualized in Figure S2

Read in associated packages
```{r, quiet=TRUE}
library(reshape2)
library(readr)
library(dplyr)
library(tidyverse)
library(tidyr)
library(car)        # for Anova()
library(emmeans)    # for multiple comparisons!
library(multcomp)
library(multcompView)
library(ggpubr)
library(lubridate)
library(circular)
```

#Diapause induction (Summer 2022, 12L:12D)#
```{r}
diapause_induction <- read_csv("../data/diapause_mutant.csv")
m1 <- glm(cbind(n_larva,n_pupa) ~ genotype,
          family = binomial(link = 'identity'),
          data = diapause_induction)

car::Anova(m1)

diapause_induction = diapause_induction %>% 
  mutate(diapause_incidence = n_larva/(n_pupa+n_larva))

summary_data <- diapause_induction %>%
  group_by(genotype) %>%
  summarize(mean_diapause = ifelse(mean(diapause_incidence, na.rm = TRUE)<= 0 , 0.01, mean(diapause_incidence, na.rm = TRUE)),
            sd_diapause = sd(diapause_incidence, na.rm = TRUE),
            n = n()) %>%
  mutate(se_diapause = sd_diapause / sqrt(n),
         lower.ci = ifelse((mean_diapause - 1.96*se_diapause) <=0, 0.01, mean_diapause - 1.96*se_diapause ),
         upper.ci = ifelse((mean_diapause + 1.96*se_diapause)>1, 1, mean_diapause + 1.96*se_diapause))

summary_data

#calculate proportion of somatic mutants
71/74
n = 74
p_hat = 71/n
alpha=0.05
# Calculate the critical z-score
z = qnorm(1-alpha/2)
# Compute the CI
p_hat + c(-1,1)*z*sqrt(p_hat*(1-p_hat)/n)
prop.test(x=c(71,0), n=c(74,59))

figureS2a <- ggplot(summary_data, aes(x = genotype, y = mean_diapause, fill = genotype, color = genotype)) +
  geom_col(position = position_dodge(), width = 0.5, alpha = 0.5) +
  geom_errorbar(aes(ymin = lower.ci, ymax = upper.ci), 
                width = 0.2, position = position_dodge(0.7), color = 'black') +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(y = "Diapause Incidence (%)", fill = "Genotype") +
  theme_bw() +
  theme(text = element_text(size = 14, family = "sans", color = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'top') +
   scale_x_discrete(breaks = c("wt", "mutant"), name = 'Genotype',
                    limits = c('wt', 'mutant'),
  labels = c('wt' = expression(italic(per)^"+"*"/"*italic(W)),
                                 'mutant' = expression(italic(per)^"-"*"/"*italic(W)))) +
  scale_fill_manual(values = c("mutant" = '#00CD00',
                                "wt" = 'darkorchid'), name = '',
                      labels = c('wt' = expression(italic(per)^"+"*"/"*italic(W)),
                                 'mutant' = expression(italic(per)^"-"*"/"*italic(W))),
                    breaks = c('wt', 'mutant')) +
  scale_color_manual(values = c("mutant" = '#00CD00',
                                "wt" = 'darkorchid'), name = '',
                      labels = c('wt' = expression(italic(per)^"+"*"/"*italic(W)),
                                 'mutant' = expression(italic(per)^"-"*"/"*italic(W))),
                     breaks = c('wt', 'mutant'))
  
figureS2a

```


#Eclosion Timing (Summer 2022, 16L:08D)#
Lights-On: 3:00-19:00
Lights-Off: 19:00-3:00

Read in data and convert time from AZT H:M to a workable format
```{r}
ecl_a <- read_csv("../data/eclosion_mutant.csv")

#convert time from ET to AZT
temp_min = lubridate::minute(ecl_a$azt)/60*100
temp_hour = lubridate::hour(ecl_a$azt)
add_leading_zeros <- function(character_vector){
  for (i in 1:length(character_vector)){
    if (is.na(character_vector[i])){
      next
    }
    if (nchar(character_vector[i]) == 2){
      next
    }
    if (nchar(character_vector[i]) == 1){
      character_vector[i] = paste0(0, character_vector[i])
      next
    }
  }
  return (character_vector)
}
ecl_a$azt = paste0(add_leading_zeros(temp_hour), ".", add_leading_zeros(temp_min)) %>% as.numeric()
ecl_a$azt[which(ecl_a$azt == 24)] <- 0 #correct azt 24 to azt0 :)
ecl_a$azt = floor(ecl_a$azt)
ecl_a$azt = circular(ecl_a$azt, units = 'hours', template='clock24') 

watson.two.test(ecl_a$azt[which(ecl_a$strain == 'be' & ecl_a$genotype == 'wt')],
                ecl_a$azt[which(ecl_a$strain == 'be' & ecl_a$genotype == 'pX4f63')])

ecl_a = ecl_a %>% group_by(genotype, family, sex) %>% filter(n() > 5) %>% filter(sex == 'female')

```

Prepare comparisons: (strain * genotype)
Comparison: B33 pX4f63 vs. BE wt
```{r}
res = ecl_a %>% group_by(strain, genotype) %>%
  summarize(
    n = n(),
    eclose.mu = ifelse(round(mean.circular(azt, na.rm = TRUE), 1)<0,
                       round(mean.circular(azt, na.rm = TRUE), 1)+24,
                       round(mean.circular(azt, na.rm = TRUE), 1)),
    eclose.sd = round(sd.circular(azt, na.rm = TRUE), 1),
    rayleigh.stat = round(rayleigh.test(azt)[[1]], 3),
    rayleigh.pval = round(rayleigh.test(azt)[[2]], 3)
  ) 
res$p.adj = round(p.adjust(res$rayleigh.pval, method='BH'),2)

res

```

plot eclosion mutant vs. wildtype
```{r}
figureS2b = ggplot(ecl_a %>% filter(strain == 'be', genotype == 'pX4f63' | genotype == 'wt'),
       aes(x=azt, color=genotype, fill=genotype)) +
  geom_histogram(alpha = 0.5, position='dodge', bins=24) +
  ylab("Count") +
  theme_bw() +
  theme(text = element_text(size = 14, family = "sans", color = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'top') +
  scale_x_continuous(breaks = seq(0,20,by = 4), limits = c(0,24), name = 'Zeitgeber Time (hours)', expand = c(0, 0)) + 
  scale_y_continuous(breaks = seq(0,12,by = 4), limits = c(-0.5,13), name = 'Number of adults eclosed', expand = c(0, 0)) + 
  scale_fill_manual(values = c("pX4f63" = "#00CD00",
                                "wt" = "darkorchid"),
                     labels = c('wt' = expression(italic(per)^"+"*"/"*italic(W)),
                                 'pX4f63' = expression(italic(per)^"-"*"/"*italic(W))),
                      breaks = c('wt', 'pX4f63'), name = '') +
  scale_color_manual(values = c("pX4f63" = '#00CD00',
                                "wt" = "darkorchid")) + guides(color = 'none') + 
  annotate("rect", xmin=0, xmax=16, ymin=-0.5, ymax=0, alpha=0.5, fill="yellow", linewidth = 1.5) +
  annotate("rect", xmin=16, xmax=24, ymin=-0.5, ymax=0, alpha=0.5, fill="black") + 
  geom_hline(yintercept = 0, linewidth = 0.15)

figureS2b
```

Combine figures and save
```{r}
figureS2 = ggarrange(figureS2a, figureS2b, common.legend = TRUE, ncol = 2)
ggsave("../plots/figureS2.png", figureS2, width = 6.25, height = 3.63, dpi = 600, units=c('in'), bg='white')

```

