---
title: "figure1"
output: html_notebook
---

Analyze ecotype x photoperiod variation in both seasonal (diapause incidence, PDD time) AND daily (activity phase, activity period length) of #O. nubilalis#, as visualized in Figure 1

Read in associated packages
```{r, quiet=TRUE}
library('car')
library('emmeans')
library('ggplot2')
library('ggpubr')
library('readr')
library('tidyverse')

```


p(diapause) = ecotype x photoperiod (Figure 1A)
```{r}
ecb_dia <- read_csv("../data/diapause_incidence_202108.csv", 
                    col_types = cols(Date = col_date(format = "%m/%d/%Y"), 
                                     Strain = col_factor(levels = c("UZ", 
                                                                    "BE",
                                                                    "NH")))) %>%
  mutate(diapause_inc = 1-Pupae/Total, Date = as.factor(Date)) %>%
  mutate(adj_diapause_inc = 1-adj_Pupae/adj_Total)

#remove NH population from analysis and model
ecb_dia_filt <- ecb_dia  %>% filter(Photoperiod != 14.9) %>%
  filter(Photoperiod != 15.0) %>%
  filter(Strain != "NH") %>%  droplevels()

ecb_dia_filt$Date <- as.factor(ecb_dia_filt$Date)

#Remove 2021-08-25 outlier; use ecb_dia_filt for analysis
ecb_dia_filt <- ecb_dia_filt %>% filter(Date != '2021-08-25') %>% droplevels()

m1 <- glm(cbind(adj_Total-adj_Pupae, adj_Pupae) ~ Photoperiod*Strain,
          family = binomial(link = 'logit'), data = ecb_dia_filt)
summary(m1)
car::Anova(m1)
summary(m1)$coef[1,1]/summary(m1)$coef[2,1]*-1 #CPP

#Create dataframe to use for predictions
newdata = with(ecb_dia_filt, expand.grid(Photoperiod = seq(14.25, 16.5, by = 0.01),
                                         Strain=unique(Strain)))
m1_pred <- predict(m1, newdata,
                   type = 'link', se = TRUE)
m1_pred <- cbind(newdata, m1_pred)

m1_pred <- within(m1_pred, {
  prob_pred <- plogis(fit) %>% round(3)
  LB <- plogis(fit - 1.96*se.fit) %>% round(3)
  UB <- plogis(fit + 1.96*se.fit) %>% round(3)
  })

m1Pred_ecbRaw <- left_join(m1_pred, ecb_dia_filt)

figure1a <-  ggplot(data = m1Pred_ecbRaw %>% filter(Strain != "NH"),
                    aes(x = Photoperiod, y=1-adj_Pupae/adj_Total, color = Strain)) +
  geom_point(position = position_jitter(width = 0.02, height = 0.02), size = 2, alpha = 0.5) + 
  geom_line(aes(x = Photoperiod, y = prob_pred), size = 0.75) +
  theme_bw() + ylab('Diapause Incidence') + xlab('Photoperiod (hours)') +
  geom_segment(x=15.05,xend=15.05,y=0.0,yend=0.5, linetype="dashed", color = "black", size = 0.5) + #UZ
  geom_segment(x=14.25,xend=15.05,y=0.5,yend=0.5, linetype="dashed", color = "black", size = 0.5) + #ALL
  geom_segment(x=14.91,xend=14.91,y=0.0,yend=0.5, linetype="dashed", color = "black", size = 0.5) + #BE
  geom_ribbon(aes(ymin = LB, ymax = UB, x=Photoperiod, fill = Strain), alpha = 0.2, colour = NA) +
  scale_y_continuous(breaks = c(0.0, 0.25, 0.50, 0.75, 1.0),
                     expand = c(0,0.00), limits = c(-0.02,1.02 ),
                     labels = scales::percent_format()) +
  scale_x_continuous(breaks = c(14.25, 15.00, 15.75, 16.50),
                     expand = c(0.0,0.0),
                     limits = c(14.17, 16.58)) +
  theme_bw() + 
  theme(text = element_text(size = 14, family = "sans", color = 'black'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), legend.position = "top") +
  scale_color_manual(values = c("UZ" = "#F8766D", "BE" = "#00BFC4"),
                     labels = c("UZ" = "Univoltine",
                                "BE" = "Bivoltine"),
                     name = "") +
  guides(fill = 'none')

figure1a

```

p(PDD) = ecotype x photoperiod (Figure 1B)
```{r}
pdd_df <- read_csv("../data/diapause_pdd_202108.csv", 
                    col_types = cols(Date = col_date(format = "%m/%d/%Y"), 
                                    Photoperiod = col_number(), 
                                    Strain = col_factor(levels = c("UZ","BE", "NH")),
                                    Transfer = col_date(format = "%m/%d/%Y"))) %>% 
  filter(Strain != "NH") %>%
  droplevels()

pdd_df_filt <- pdd_df %>% mutate(Location = NULL, Date = NULL, Transfer = NULL) %>%
  reshape2::melt(id = c("Strain", "Photoperiod", "Replicate", "Total"),
       measure = colnames(pdd_df)[8:ncol(pdd_df)],
       na.rm = FALSE, variable.name = "Day", value.name = "Pupae")

pdd_df_filt$Replicate <- as.factor(pdd_df_filt$Replicate)
pdd_df_filt$Day <- as.numeric(as.character(pdd_df_filt$Day))
pdd_df_filt$Pupae <- as.numeric(as.character(pdd_df_filt$Pupae))
pdd_df_filt$Total <- as.numeric(as.character(pdd_df_filt$Total))

pdd_df_filt <- pdd_df_filt %>% filter(Total != 0) %>% na.omit()
pdd_df_filt$Strain = as.factor(pdd_df_filt$Strain)
pdd_df_filt$Strain = relevel(pdd_df_filt$Strain, ref = 'UZ')

m1_pdd <- glm(cbind(Pupae, Total-Pupae) ~ Day*Photoperiod + Photoperiod*Strain + Day*Strain,
              family = binomial(link = 'logit'), data = pdd_df_filt)
car::Anova(m1_pdd)

pdd_new <- with(pdd_df_filt, expand.grid(Photoperiod = seq(14.25, 15.25, by = 0.5),
                                          Day=seq(0,61,by=1), Strain = unique(Strain)))

pdd_pred <- predict.glm(m1_pdd, pdd_new, type="response") %>% round(digits = 3)
df_pred <- cbind(pdd_new, "Pupation" = pdd_pred)
df_pred$Photoperiod <- as.factor(df_pred$Photoperiod)

pdd_pred <- cbind(pdd_new, predict.glm(m1_pdd,
                                           pdd_new,
                                           type="response", se = TRUE))
pdd_pred <- within(pdd_pred, {
  prob_pred <- fit %>% round(3)
  LB <- ifelse((fit - 3*qnorm(0.975)*se.fit %>% round(3))<0, 0,fit -  3*qnorm(0.975)*se.fit %>% round(3) )
  UB <- ifelse((fit + 3*qnorm(0.975)*se.fit %>% round(3))>1, 1,fit + 3*qnorm(0.975)*se.fit %>% round(3) )
  })

df_pred <- rbind(filter(pdd_pred, Strain == "UZ"),
                 filter(pdd_pred, Strain == "BE"))

figure1b <- ggplot(data = df_pred,
                          aes(x=Day, y = prob_pred, color = Strain)) + theme_bw() +
  theme(axis.text.x = element_text(size = 14, family = 'sans')) +
  theme(text = element_text(size = 14, family = 'sans')) +
  theme(legend.position = "top", panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  ylim(0, 1) + xlim(1, 61) +
  scale_shape_manual(values=c(16, 2, 4)) +
  geom_line(data = df_pred %>% filter(Photoperiod == 14.25),
            aes(x=Day, y = prob_pred, color = Strain, linetype=as.factor(Photoperiod)), size = 0.75) +
    scale_y_continuous(breaks = c(0.0, 0.25, 0.50, 0.75, 1.0),
                     expand = c(0,0.00), limits = c(-0.02,1.02 ),
                     labels = scales::percent_format(), name = "Cumultative Pupation") +
  xlab("PDD Time (days)") +
  scale_linetype_manual(values=c(1,3,6)) + 
  geom_ribbon(data = df_pred %>% filter(Photoperiod == 14.25), 
              aes(ymin = LB, ymax = UB, x=Day, fill = interaction(Strain, Photoperiod)), 
              alpha = 0.2, colour = NA, show.legend = FALSE) +
  guides(linetype = 'none') +
  scale_color_manual(values = c("UZ" = "#F8766D", "BE" = "#00BFC4"),
                     labels = c("UZ" = paste("Univoltine"),
                                "BE" = paste("Bivoltine")),
                     name = "")

figure1b <- figure1b + geom_point(data = pdd_df_filt %>%
                                                  filter(Photoperiod == 14.25),
                                                aes(x = Day, y = Pupae/Total, color = Strain),
                                                size = 2, alpha = 0.5,
                                                position = position_jitter(width = 0.05, height = 0.05)) +
  geom_segment(x=0,xend=38.3,y=0.5,yend=0.5, linetype="dashed", color = "black", size = 0.5) +
  geom_segment(x=38.2,xend=38.2,y=0.0,yend=0.5, linetype="dashed", color = "black", size = 0.5) +
  geom_segment(x=19.6,xend=19.6,y=0,yend=0.5, linetype="dashed", color = "black", size = 0.5) 

figure1b

```


activity time = ecotype x photoperiod (Figure 1C)
```{r}
phase_df <- read_csv("../data/activity_summary.csv", 
                         col_types = cols(...1 = col_skip()),
                     show_col_types = FALSE)

phase_df$strain = as.factor(phase_df$strain)
phase_df$strain = relevel(phase_df$strain, ref = 'univoltine')

#lm
m1 = lm(peak_after_day ~ daylength+strain, data = phase_df)
car::Anova(m1)
m1_contrast = emmeans::contrast(emmeans::emmeans(m1, ~ strain, type = 'response'), method = 'pairwise')
summary(m1_contrast, infer = c(TRUE, TRUE))

figure1c = ggplot(data=phase_df,
                 aes(x=daylength, y=peak_after_day, color=strain)) +
  geom_jitter(size=2, alpha=0.5, position = position_jitter(width = 0.04, height = 0.02)) +
  geom_smooth(method='lm', se=TRUE, aes(fill = strain),alpha = 0.2, size = 0.75) +
  scale_fill_manual(values = c('univoltine' = "#F8766D", 'bivoltine' = "#00BFC4"),
                     labels = c('univoltine' = "Univoltine",
                                'bivoltine' = "Bivoltine")) +
  scale_color_manual(values = c('univoltine' = "#F8766D", 'bivoltine' = "#00BFC4"),
                     labels = c('univoltine' = "Univoltine",
                                'bivoltine' = "Bivoltine")) +
  theme_bw() +
  theme(text = element_text(size = 14, family = "sans", color = 'black'),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        legend.title = element_blank()) +
  ylab("Peak Activity (after dark)") + xlab("Photoperiod (hours)") +
  scale_x_continuous(breaks = c(12, 13, 14, 15, 16), expand = c(0.0,0), limits = c(11.9, 16.1)) +
  scale_y_continuous(breaks = c(4,6,8), expand = c(0,0), limits = c(3,9)) +
  guides(fill = 'none')

```

free-running period length = ecotype (Figure 1D)
```{r}
circ_df <- read_csv("../data/activity_circadian.csv")
circ_df$strain = as.factor(circ_df$strain)
circ_df$strain = relevel(circ_df$strain, ref = 'univoltine')

t.test(data = circ_df, period ~ strain)

figure1d = ggplot(circ_df,
                  aes(x=strain, y=period, color = strain)) +
  geom_boxplot(size = 0.5) +
  geom_point(aes(x = strain, y = period, color = strain), alpha=0.5, size=2) +
  theme_bw() +
  ylab("Period (hours)") + xlab("Population") +
  scale_color_manual(values = c(
    'bivoltine' = "#00BFC4",
    'univoltine' = "#F8766D"),
    labels = c(
      'bivoltine' = 'Bivoltine',
      'univoltine' = 'Univoltine',
    name = "")) +
  theme(text = element_text(size = 14, family = "sans", color = 'black'),
        strip.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = "top",
        legend.title = element_blank()) +
  scale_x_discrete(breaks = c('bivoltine', 'univoltine'), limits = c('bivoltine', 'univoltine'),
                   labels = c('bivoltine' = 'Bivoltine', 'univoltine' = 'Univoltine'),
                   name = "Population") +
  scale_y_continuous(breaks = seq(21, 23,  by = 1), expand = c(0,0), limits = c(20.5,23.5),
                   name = 'Period Length (hours)') +
  annotate("text", x=1.5, y=23.25, label= "**", size = 4)

```


Build plot (Figure 1)
```{r}
figure1 = ggarrange(figure1a,figure1b,
                    figure1c,figure1d,
                    ncol = 2, nrow = 2,
                    common.legend = TRUE, legend = 'top', align = 'hv')


ggsave("../plots/figure1.png", plot = figure1, width = 6, height = 6, dpi = 600, units=c('in'),
               bg='white')
```

